---
title: "p8105_hw2_pw2551"
author: "Paula Wu"
date: "10/1/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the libraries:
```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1
Since I'm going to use the file path fairly frequently later, I'd like to define it beforehand.
```{r}
file_path = "./Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

Read and clean the Mr.Trash Wheel dataset:
```{r}
wheels_df = 
  read_excel(file_path, range = "Mr. Trash Wheel!A2:N408") %>% 
  janitor::clean_names() %>%
  filter(!(is.na(dumpster))) %>% # filter out rows that don't have dumpster-specific data
  mutate(sports_balls = round(sports_balls, 0)) # round sports_balls to nearest integer
wheels_df
```
<br> Read and clean precipitation data for 2018:
```{r}
year_2018 = 
  read_excel(file_path, range = "2018 Precipitation!A2:B14") %>%  # only cells between A2:B14 are relevant
  janitor::clean_names() %>% 
  mutate(year = 2018)
year_2018
```
Read and clean precipitation data for 2019: we only got precipitation data from Jan to Jun during 2019
```{r}
year_2019 = 
  read_excel(file_path, range = "2019 Precipitation!A2:B8") %>%
  janitor::clean_names() %>% 
  mutate(year = 2019)
year_2019
```
Combine previous two data sets:
```{r}
# stack these two data sets together
two_years = 
  bind_rows(year_2018, year_2019) %>% 
  mutate(month = month.name[month])  # convert month to char variable
two_years
```
## Problem 2 ##
Read and manipulate data from "pols-month.csv"
```{r}
pols_month = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv", show_col_types = FALSE) %>% 
  separate(mon, into = c('year', 'month', 'day'), sep = "-") %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  select(-c(prez_dem, prez_gop, day))
pols_month
```




