---
title: "p8105_hw2_pw2551"
author: "Paula Wu"
date: "10/1/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the libraries:
```{r, message = FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

## Problem 1
Since I'm going to use the file path fairly frequently later, I'd like to define it beforehand.
```{r}
file_path = "./Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

Read and clean the Mr.Trash Wheel dataset:
```{r}
wheels_df = 
  read_excel(file_path, range = "Mr. Trash Wheel!A2:N408") %>% 
  janitor::clean_names() %>%
  filter(!(is.na(dumpster))) %>% # filter out rows that don't have dumpster-specific data
  mutate(sports_balls = round(sports_balls, 0)) # round sports_balls to nearest integer
knitr::kable(wheels_df[0:5,])
```
<br> Read and clean precipitation data for 2018:
```{r}
year_2018 = 
  read_excel(file_path, range = "2018 Precipitation!A2:B14") %>%  # only cells between A2:B14 are relevant
  janitor::clean_names() %>% 
  mutate(year = 2018)
knitr::kable(year_2018[0:10,])
```
Read and clean precipitation data for 2019: we only got precipitation data from Jan to Jun during 2019
```{r}
year_2019 = 
  read_excel(file_path, range = "2019 Precipitation!A2:B8") %>%
  janitor::clean_names() %>% 
  mutate(year = 2019)
knitr::kable(year_2019[0:6,])
```
Combine previous two data sets:
```{r}
# stack these two data sets together
two_years = 
  bind_rows(year_2018, year_2019) %>% 
  mutate(month = month.name[month])  # convert month to character variable
knitr::kable(list(two_years[0:9,], two_years[10:18,]), format = 'html', booktabs = TRUE)
```
## Problem 2 ##
Read and manipulate data from "pols-month.csv"
```{r}
pols = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv", show_col_types = FALSE) %>% 
  separate(mon, into = c('year', 'month', 'day'), sep = "-", convert = TRUE) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(president = recode(prez_dem, `1` = "dem", `0` = "gop")) %>% 
  #filter(prez_gop == 2)
  select(-c(prez_dem, prez_gop, day))
pols
```

```{r}
#unique(pull(pols, prez_gop))
```

<br> Read and manipulate data from "snp.csv"

```{r}
snp = 
  read_csv("./fivethirtyeight_datasets/snp.csv", show_col_types = FALSE) %>% 
  mutate(date = parse_date_time2(date, 'mdy', cutoff_2000 = 49)) %>% 
  separate(date, into = c('year', 'month', 'day'), sep = "-", convert = TRUE) %>% 
  mutate(month = month.name[month]) %>% 
  arrange(year, month) %>%
  select(-day) %>% 
  select(year, month, everything())
view(snp)
```

<br> Read and manipulate data from "unemployment.csv"
```{r}
unemploy =
  read_csv("./fivethirtyeight_datasets/unemployment.csv", show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:dec, names_to = "month", values_to = "unemploy_rate") %>% 
  mutate(year = as.integer(year)) %>% # change type to integer for `year`
  mutate(month = month.name[match(str_to_title(month), month.abb)])  #acceptable?
unemploy
```
Merge `snp` into `pols`:
```{r}
merge_1 = left_join(pols, snp, by = c('year', 'month'))
merge_1
```


## Problem 3##
Read data, clean variable names and childs' first names
```{r}
baby_names = 
  read_csv("./Popular_Baby_Names.csv",show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  mutate(childs_first_name = str_to_title(childs_first_name))
```

Check for ethnicity names
```{r}
# check different names of variable `ethnicity`
unique(pull(baby_names, ethnicity))
```
It's clear that people change the ethnicity encoding overtime: "BLACK NON HISPANIC" used to be coded as "BLACK NON HISP", "WHITE NON HISPANIC" used to be coded as "WHITE NON HISP", and etc. <br>
I will rename the variable `ethnicity`, and make sure it is consistent across the data frame.
```{r}
baby_names = 
  baby_names %>% 
  mutate(ethnicity = replace(ethnicity, grepl('BLACK', ethnicity), 'BLACK NON HISPANIC')) %>% 
  mutate(ethnicity = replace(ethnicity, grepl('WHITE', ethnicity), 'WHITE NON HISPANIC')) %>%
  mutate(ethnicity = replace(ethnicity, grepl('ASIAN', ethnicity), 'ASIAN AND PACIFIC ISLANDER'))

# check consistency
unique(pull(baby_names, ethnicity))
```
There are only four groups, which means I successfully tidy values in `ethnicity` variable. <br>
And then remove duplicate rows:
```{r}
baby_names = 
  baby_names %>% 
  distinct(year_of_birth, gender, ethnicity, childs_first_name, count, rank, .keep_all = TRUE)
baby_names
```
<br> How the name **Olivia** change overtime
```{r}
olivias = 
  baby_names %>% 
  filter(childs_first_name == "Olivia") %>% 
  select(-c(childs_first_name, gender, count)) %>% 
  pivot_wider(names_from = "year_of_birth", values_from = "rank")
olivias
```
Popular boys names
```{r}
# search for the most popular names for boys
boys = 
  baby_names %>% 
  filter(rank == 1 & gender == "MALE") %>% 
  select(-c(gender, rank, count)) %>% 
  pivot_wider(names_from = "year_of_birth", values_from = "childs_first_name")
boys
```
<br> Plotting
```{r}
male_white_16 = 
  baby_names %>% 
  filter(year_of_birth == 2016 & ethnicity == "WHITE NON HISPANIC" & gender == "MALE")
ggplot(male_white_16, aes(x = rank, y = count)) + 
  geom_point() + 
  scale_x_reverse() +
  ggtitle("Count and Rank of a Child's First Name") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "# of Children with a Name", x = "Rank of a Name")
```


