---
title: "p8105_hw2_pw2551"
author: "Paula Wu"
date: "10/1/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the libraries:
```{r, message = FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

## Problem 1
Since I'm going to use the file path fairly frequently later, I'd like to define it beforehand.
```{r}
file_path = "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx"
```

Read and clean the Mr.Trash Wheel dataset:
```{r, warning=FALSE}
wheels_df = 
  read_excel(file_path, range = "Mr. Trash Wheel!A2:N535") %>% 
  janitor::clean_names() %>%
  filter(!(is.na(dumpster))) %>% # filter out rows that don't have dumpster-specific data
  filter(!is.na(as.numeric(as.character(dumpster)))) %>% # filter out the last row: the grand total
  mutate(sports_balls = round(sports_balls, 0)) # round sports_balls to nearest integer
knitr::kable(wheels_df[0:5,])
```
Read and clean precipitation data for 2018:
```{r}
year_2018 = 
  read_excel(file_path, range = "2018 Precipitation!A2:B14") %>%  # only cells between A2:B14 are relevant
  janitor::clean_names() %>% 
  mutate(year = 2018)
knitr::kable(year_2018[0:10,])
```
Read and clean precipitation data for 2019:
```{r}
year_2019 = 
  read_excel(file_path, range = "2019 Precipitation!A2:B14") %>%  # only cells between A2:B14 are relevant
  janitor::clean_names() %>% 
  mutate(year = 2019)
knitr::kable(year_2019[0:10,])
```
Combine previous two data sets:
```{r}
# stack these two data sets together
two_years = 
  bind_rows(year_2018, year_2019) %>% 
  mutate(month = month.name[month])  # convert month to character variable
knitr::kable(two_years[8:16,])  # arbitrarily choose to show the data from Aug.2018 - Apr.2019
```

```{r}
# total precipitation 2018
tot_prcp_2018 = sum(pull(year_2018, total))

# median of sports balls in 2019
data_2019 = filter(wheels_df, year == 2019)
md = median(pull(data_2019, sports_balls))
```

#### Data description <br>
After tidying and merging data extracted from Mr. Trash Wheel data set, I have two data frames in the end: `wheels_df` and `two_years`. <br>
-- The `wheels_df` data frame has `r nrow(wheels_df)` rows and `r ncol(wheels_df)` columns, after omitting non-dumpter specific data. Among those features, I think `weight_tons` and `volume_cubic_yards` are important since they gave us a direct view of how much trash Mr. Trash Wheel has collected during a day. The mean of `weight_tons` is `r round(mean(pull(wheels_df, weight_tons)),2)` tons, and the mean of `volume_cubic_yards` is `r round(mean(pull(wheels_df, volume_cubic_yards)),2)` cubic yards - an impressive amount of trash has been collected by Mr. Trash Wheel per day. <br>
-- The `two_years` data frame was created by merging data from "2018 Precipitation" and "2019 Precipitation", and it has `r nrow(two_years)` rows and `r ncol(two_years)` columns. The variable `total`, which means total precipitation during that month, is quite important. The mean of it is `r round(mean(pull(two_years, total)),2)`. <br>
-- The total precipitation during 2018 is `r tot_prcp_2018`, and the number of sports balls in 2019 is `r md`. <br>


## Problem 2 ##
*Note: for concision purpose, I set `show_col_types` equals to FALSE every time I read a csv file* <br>
<br> Read and manipulate data from "pols-month.csv"
```{r}
pols = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv", show_col_types = FALSE) %>% 
  separate(mon, into = c('year', 'month', 'day'), sep = "-", convert = TRUE) %>% 
  mutate(month = month.name[month])
```
How I determine the values for `president`: <br>
- At first, I assume `prez_dem` and `prez_gop` are both binary variables. However, `unique(pull(pols, prez_gop))` shows the following result: `r unique(pull(pols, prez_gop))`. There's an extra **2**, which I assume it was due to entry error. I pulled out entries that has a value of 2 for the variable `prez_gop`, and it shows the following results:
```{r}
filter(pols, prez_gop == 2) %>% 
  select(prez_gop, prez_dem, year, month)  # only select a few lines for concise output
```
The table confirms my assumption that the 2's are probably due to data entry error, and it has the same meaning as 1's. Due to the inaccuracy in `prez_gop` column, I decide to use `prez_dem` to code for my `president` column.
```{r}
pols = 
  pols %>% 
  mutate(president = recode(prez_dem, `1` = "dem", `0` = "gop")) %>% 
  select(-c(prez_dem, prez_gop, day))
knitr::kable(pols[0:10,])
```

<br> Read and manipulate data from "snp.csv"
```{r}
snp = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv", show_col_types = FALSE) %>% 
  mutate(date = parse_date_time2(date, 'mdy', cutoff_2000 = 49)) %>% 
  separate(date, into = c('year', 'month', 'day'), sep = "-", convert = TRUE) %>%
  arrange(year, month) %>% 
  mutate(month = month.name[month]) %>% 
  select(-day) %>% 
  select(year, month, everything())
knitr::kable(snp[0:8,])
```
<br> Read and manipulate data from "unemployment.csv"
```{r}
unemploy =
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv", show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:dec, names_to = "month", values_to = "unemploy_rate") %>% 
  filter(!(is.na(unemploy_rate))) %>% # omit NULL data after Jun 2015 
  mutate(year = as.integer(year)) %>% # change type to integer for `year`
  mutate(month = month.name[match(str_to_title(month), month.abb)])  # for consistent month names
knitr::kable(unemploy[0:10,])
```
<br>Merge `snp` into `pols`:
```{r}
merge_1 = left_join(pols, snp, by = c('year', 'month'))
merge_1 = left_join(merge_1, unemploy, by = c('year', 'month'))
```
Just for more straightforward presentation, I rearranged the data set in the way that I think could help interpretation.  
```{r}
merge_1 = 
  merge_1 %>% 
  select(year, month, president, close, unemploy_rate, everything())
knitr::kable(merge_1[11:21,])  # data entries displayed here are arbitrarily selected
```
#### Data description <br>
* `pols` is the largest data set among these three, both in dimension (`r dim(pols)`) and in year range (from Jan.1947 to Jun.2015). After data tidying, it mainly contains year, month, the number of either GOP or Democratic governors/ senators/ representatives on a specific day, and the current president's party. <br>
* `snp` has a dimension of `r dim(snp)`. After data tidying and manipulation, it contains year, month, and the closing value of the S&P stock. <br>
* `unemploy` has a dimension of `r dim(unemploy)`. It contains year, month, and unemployment rate of the associate month. <br>
* All data sets have different year range. After merging, the final data set has a dimension of `r dim(merge_1)`. The year ranges from Jan.1947 to Jun.2015, which is the same as `pols` since we used `left_join()`. As I did above, I rearranged the data set and put five variables that I think is important upfront: `r variable.names(merge_1)[1:5]`.<br>
* Please be noted that there are many NAs in both `close` and `unemploy_rate` columns, this is because data in `snp` and `umemploy` aren't consistently populated until Jan.1950 and Jan.1948, respectively. Also, `snp` has a data entry for Jul.2015 but `pols` doesn't, and this entry is excluded due to the use of `left_join()`. <br>

## Problem 3##
Read data, clean variable names, gender, child's first names
```{r}
baby_names = 
  read_csv("./data/Popular_Baby_Names.csv",show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  mutate(childs_first_name = str_to_title(childs_first_name)) %>% 
  mutate(gender = str_to_lower(gender))
```
I set all values in `gender` to lower cases for a better-looking data set. And I reformatted `childs_first_name` because there are several entries where first names are all capitalized.<br>
<br> Check for different ethnicity values
```{r}
# check different values in the variable `ethnicity`
unique(pull(baby_names, ethnicity))
```
It's clear that people change the ethnicity encoding overtime: "BLACK NON HISPANIC" used to be coded as "BLACK NON HISP", "WHITE NON HISPANIC" used to be coded as "WHITE NON HISP", and etc. <br>
I will reformat values in `ethnicity`, and make sure it is consistent across the data frame.
```{r}
# use find and replace
baby_names = 
  baby_names %>% 
  mutate(ethnicity = replace(ethnicity, grepl('BLACK', ethnicity), 'Black non hispanic')) %>% 
  mutate(ethnicity = replace(ethnicity, grepl('WHITE', ethnicity), 'White non hispanic')) %>%
  mutate(ethnicity = replace(ethnicity, grepl('ASIAN', ethnicity), 'Asian and pacific islander')) %>% 
  mutate(ethnicity = replace(ethnicity, grepl('HIS', ethnicity), 'Hispanic'))

# check consistency
unique(pull(baby_names, ethnicity))
```
There are only four unique values in `ethnicity`, which means I successfully tidy the data in this column. <br>
And then remove duplicated rows:
```{r}
baby_names = 
  baby_names %>% 
  distinct(year_of_birth, gender, ethnicity, childs_first_name, count, rank, .keep_all = TRUE)
knitr::kable(baby_names[0:10,])
```
<br> How the name *Olivia* change overtime
```{r}
olivias = 
  baby_names %>% 
  filter(childs_first_name == "Olivia") %>% 
  select(-c(childs_first_name, gender, count)) %>%  # get rid of not-so-useful columns
  pivot_wider(names_from = "year_of_birth", values_from = "rank") %>% 
  select(ethnicity, sort(names(.)))
knitr::kable(olivias)
```
<br>Popular boys names across time
```{r}
# search for the most popular names for boys
boys = 
  baby_names %>% 
  filter(rank == 1 & gender == "male") %>% 
  select(-c(gender, rank, count)) %>%  # get rid of not-so-useful columns 
  pivot_wider(names_from = "year_of_birth", values_from = "childs_first_name") %>% 
  select(ethnicity, sort(names(.)))
knitr::kable(boys)
```
<br> Plotting
```{r}
male_white_16 = 
  baby_names %>% 
  filter(year_of_birth == 2016 & ethnicity == "White non hispanic" & gender == "male")
ggplot(male_white_16, aes(x = rank, y = count)) + 
  geom_point() + 
  ggtitle("Count and Rank of a Child's First Name") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "# of Children with a Name", x = "Rank of a Name", 
       caption = "data from male white non hispanic children born in 2016")
```


